---
title: "Capstone Project - Pump it up"
author: "Tadaaki Sun"
output: pdf_document
---

## Capstone Project - Pump it up

Load Libraries:
```{r}
# Load libraries
library(ggplot2)
library(rpart)
library(rpart.plot)
library(caTools)
library(ROCR)
```

# Uploading Data

Import train values and labels
```{r}
train_values <- read.csv("train_value.csv")
train_labels <- read.csv("train_label.csv")
```

Merge data frames
```{r}
train <- merge(train_labels, train_values)
```

Optional cleaning environment
```{r}
rm(train_values)
rm(train_labels)
```

Preliminary Data Analysis
```{r}
summary(train)
```

Split
```{r}
library(caret)
set.seed(3456)
Index <- createDataPartition(train$status_group, p = 0.7, 
                                  list = FALSE, 
                                  times = 1)
head(Index)

train <- train[ Index,]
test  <- train[-Index,]

#split = sample.split(train$status_group, SplitRatio = 2/3, group=NULL)
#train = subset(train, split == TRUE)
#test = subset(train, split == FALSE)
#nrow(train)
#nrow(test)
```

Merging "functional needs repair" with "non functional" 
```{r}
train$status_group <- replace(train$status_group, train$status_group == "functional needs repair", "non functional")
train$status_group <- factor(train$status_group)
test$status_group <- replace(test$status_group, test$status_group == "functional needs repair", "non functional")
test$status_group <- factor(test$status_group)
summary(train$status_group)
summary(test$status_group)
```

Calculating the baseline performance
```{r}
22582/41581 # Baseline Performance = functional water points/total = 0.5430846
```

### Preliminary Exploration 

region 
```{r}
str(train$region) #21 levels is good amount, keep.

qplot(region, data=train, geom="bar", fill=status_group) + 
  theme(legend.position = "top")
```

public_meeting - 3000 missing. Decide what to do.
```{r}
str(train$public_meeting) #keep

qplot(public_meeting, data=train, geom="bar", fill=status_group) + 
  theme(legend.position = "top")
```

permit - Watch the 3000 missing values. Otherwise good amount.
```{r}
str(train$permit)

qplot(permit, data=train, geom="bar", fill=status_group) + 
  theme(legend.position = "top")
```

basin - Approximately 12000 out of 59400.
```{r}
summary(train$basin)

qplot(basin, data=train, geom="bar", fill=status_group) + 
  theme(legend.position = "top")
```

payment_type
```{r}
summary(train$payment_type)

qplot(payment_type, data=train, geom="bar", fill=status_group) + 
  theme(legend.position = "top")

train$paytype0 <- 0
train$paytype0[train$payment_type=="never pay"]<-1
table(train$paytype0)

test$paytype0 <- 0
test$paytype0[test$payment_type=="never pay"]<-1

qplot(paytype0, data=train, geom="bar", fill=status_group) + 
  theme(legend.position = "top")
```  

management_group
```{r}
summary(train$management_group)

qplot(management_group, data=train, geom="bar", fill=status_group) + 
  theme(legend.position = "top")

train$management0 <- 0
train$management0[train$management_group=="user-group"]<-1
table(train$management0)

test$management0 <- 0
test$management0[test$management_group=="user-group"]<-1

qplot(management0, data=train, geom="bar", fill=status_group) + 
  theme(legend.position = "top")
```

extraction_type_group
```{r}
str(train$extraction_type_group)

qplot(extraction_type_group, data=train, geom="bar", fill=status_group) + 
  theme(legend.position = "top")

prop.table((table(train$extraction_type_group)))
train$extraction_type_group[train$extraction_type_group=="rope pump"] <- "other"
train$extraction_type_group[train$extraction_type_group=="wind-powered"] <- "other"
train$extraction_type_group<-droplevels(train$extraction_type_group)
prop.table((table(train$extraction_type_group)))

test$extraction_type_group[test$extraction_type_group=="rope pump"] <- "other"
test$extraction_type_group[test$extraction_type_group=="wind-powered"] <- "other"
test$extraction_type_group<-droplevels(test$extraction_type_group)

qplot(extraction_type_group, data=train, geom="bar", fill=status_group) + 
  theme(legend.position = "top")

train$exttype0 <- 0
train$exttype0[train$extraction_type_group=="gravity"]<-1
table(train$exttype0)

test$exttype0 <- 0
test$exttype0[test$extraction_type_group=="gravity"]<-1

qplot(exttype0, data=train, geom="bar", fill=status_group) + 
  theme(legend.position = "top")

```

waterpoint_type
```{r}
str(train$waterpoint_type)

qplot(waterpoint_type, data=train, geom="bar", fill=status_group) + 
  theme(legend.position = "top")

prop.table(table(train$waterpoint_type))
train$waterpoint_type[train$waterpoint_type=="cattle trough"] <- "other"
train$waterpoint_type[train$waterpoint_type=="dam"] <- "other"
train$waterpoint_type[train$waterpoint_type=="improved spring"] <- "other"
train$waterpoint_type<-droplevels(train$waterpoint_type)
prop.table((table(train$waterpoint_type)))

test$waterpoint_type[test$waterpoint_type=="cattle trough"] <- "other"
test$waterpoint_type[test$waterpoint_type=="dam"] <- "other"
test$waterpoint_type[test$waterpoint_type=="improved spring"] <- "other"
test$waterpoint_type<-droplevels(test$waterpoint_type)

qplot(waterpoint_type, data=train, geom="bar", fill=status_group) + 
  theme(legend.position = "top")

#train$wptype0 <- 0
#train$wptype0[train$waterpoint_type=="communal standpipe multiple"]<-1
#train$wptype0[train$waterpoint_type=="other"]<-1
#table(train$wptype0)

#test$wptype0 <- 0
#test$wptype0[test$waterpoint_type=="communal standpipe multiple"]<-1
#test$wptype0[test$waterpoint_type=="other"]<-1

#qplot(wptype0, data=train, geom="bar", fill=status_group) + 
 # theme(legend.position = "top")

```

gps_height
```{r}
hist(train$gps_height)
nrow(train[train$gps_height==0,])

ggplot(train, aes(x = gps_height)) +
  geom_histogram(bins = 20) + 
  facet_grid( ~ status_group)

train$gps0 <- 0
train$gps0[train$gps_height==0]<-1
table(train$gps0)

test$gps0 <- 0
test$gps0[test$gps_height==0]<-1

#gps_height - how many values were missing?

hist(train$gps_height)
hist(train$gps_height[train$gps0 ==1])

qplot(gps0, data=train, geom="bar", fill=status_group) + 
  theme(legend.position = "top")

```

Quantity
```{r} 
table(train$quantity,train$status_group)
prop.table(table(train$quantity, train$status_group), margin = 1)

# Bar plot for quantity - significant reason for failure (dry)
qplot(quantity, data=train, geom="bar", fill=status_group) + 
  theme(legend.position = "top")
```

water_quality
```{r}  
# Bar plot for water_quality - significant reason for failure (salty and unknown)
str(train$water_quality)

qplot(water_quality, data=train, geom="bar", fill=status_group) + 
  theme(legend.position = "top")

train$wq0 <- 0
train$wq0[train$water_quality=="soft"]<-1
table(train$wq0)

test$wq0 <- 0
test$wq0[test$water_quality=="soft"]<-1

qplot(wq0, data=train, geom="bar", fill=status_group) + 
  theme(legend.position = "top")
```

scheme_management
```{r} 
# Bar plot for waterpoint_type - significant reason for failure (other)
str(train$scheme_management)

qplot(scheme_management, data=train, geom="bar", fill=status_group) + 
  theme(legend.position = "top") + 
  theme(axis.text.x=element_text(angle = -20, hjust = 0))

train$schman0 <- 0
train$schman0[train$scheme_management=="VWC"]<-1
table(train$schman0)

test$schman0 <- 0
test$schman0[test$scheme_management=="VWC"]<-1

qplot(schman0, data=train, geom="bar", fill=status_group) + 
  theme(legend.position = "top") + 
  theme(axis.text.x=element_text(angle = -20, hjust = 0))

```

source
```{r} 
# Bar plot for source - no significant reason for failure
str(train$source)

qplot(source, data=train, geom="bar", fill=status_group) + 
  theme(legend.position = "top") + 
  theme(axis.text.x=element_text(angle = -20, hjust = 0))

prop.table(table(train$source))
train$source[train$source=="hand dtw"] <- "other"
train$source[train$source=="dam"] <- "other"
train$source[train$source=="lake"] <- "other"
train$source[train$source=="rainwater harvesting"] <- "other"
train$source[train$source=="unknown"] <- "other"
train$source<-droplevels(train$source)
prop.table((table(train$source)))

test$source[test$source=="hand dtw"] <- "other"
test$source[test$source=="dam"] <- "other"
test$source[test$source=="lake"] <- "other"
test$source[test$source=="rainwater harvesting"] <- "other"
test$source[test$source=="unknown"] <- "other"
test$source<-droplevels(test$source)

qplot(source, data=train, geom="bar", fill=status_group) + 
  theme(legend.position = "top") + 
  theme(axis.text.x=element_text(angle = -20, hjust = 0))
```

###Exploratory Data Analysis
#Method 1 - Logistic Regression
```{r}
PumpLog <- glm(formula = status_group ~ region + public_meeting + basin + permit + paytype0 + exttype0 + waterpoint_type + gps0 + quantity + schman0 + source,
               family = "binomial"(link="logit"), data = train) 
summary(PumpLog)
coef(summary(PumpLog))
```

#Finalized Model
```{r}
predTrain <- predict(PumpLog,train, type="response")
summary(predTrain)

tapply(predTrain, train$status_group, mean, na.rm=TRUE) # 0.319 Functional, 0.62 Non Functional 
table(train$status_group, predTrain>0.5) #This model perdicts with 72.82% accuracy

16230+3134
16230/19364 # specifity = 0.838
9712+6431
9712/16143# sensitivity = 0.602
```

# Prediction
```{r}
predTest = predict(PumpLog, type="response", newdata = test)
summary(predTest)
tapply(predTest, test$status_group,mean,na.rm=TRUE) # 0.32 Functional, 0.62 Non-functional
table(test$status_group, predTest>0.5) #This model perdicts with 69.54% accuracy
```